OptiFit: a fast method for fitting amplicon sequences to existing OTUs

### 2021-02-15 paper/paper.Rmd
Re-make plots with RDP v18
Capitalization

- **TODO: brief description of OptiFit algorithm**
- **TODO: open & closed reference modes**
- **TODO: mcc score; only considers query seqs (when printref=f)**

@@

- TODO: separate paragraphs for closed & open reference clustering. weave in vsearch within those paragraphs. mention median mcc values.
- TODO: avoid slightly/much -- use numbers, e.g. X% better/worse.
- TODO: put the comparison at the beginning of the sentence, so you don't have to know what it is at the end.

To evaluate reference-based clustering with independent databases, we fit each
dataset to reference OTUs generated by _de novo_ clustering the Greengenes
database (v13_8_99), Silva non-redundant database (v132), and the Ribosomal
Database Project (RDP; v16). In open reference mode, OTU quality was similar
between fitting the datasets to reference OTUs with OptiFit and clustering the
datasets _de novo_ with OptiClust. This held true for all datasets and reference
databases. However, in closed reference mode, OTU quality was slightly worse
when fitting to greengenes and silva, and much worse when fitting to RDP as
compared to OptiClust. No more than half of query sequences were fit to
reference OTUs in closed reference mode across any dataset/database combination.
**TODO: specify exact numbers for fraction mapped.**
OptiFit was able to fit more query sequences to reference OTUs created with the
greengenes and silva databases than with RDP. In terms of runtime,
closed reference OptiFit outperformed OptiClust, while OptiClust outperformed
open reference OptiFit.

@@

- **TODO:** for these data, we don't see a compelling reason to use reference-based clustering over _de novo_. it was supposed to speed things up. the reason to do reference-based is if you like the ref OTUs -- e.g. for ML or downstream tools e.g. PiCrust?
- **TODO:** highlight difference between what we're doing and what previously was done.
  others use single ref seq to define an otu, while we use all the ref & query seqs to define an otu. highlight why ours is better than previous methods.

>>>>

- **TODO: brief description of OptiFit algorithm**
- **TODO: open & closed reference modes**
- **TODO: MCC score; only considers query sequences (when printref=f)**

@@

- TODO: separate paragraphs for closed & open reference clustering. weave in vsearch within those paragraphs. mention median MCC values.
- TODO: avoid slightly/much -- use numbers, e.g. X% better/worse.
- TODO: put the comparison at the beginning of the sentence, so you don't have to know what it is at the end.

To evaluate reference-based clustering with independent databases, we fit each
dataset to reference OTUs generated by _de novo_ clustering the Greengenes
database (v13_8_99), Silva non-redundant database (v132), and the Ribosomal
Database Project (RDP; v16). In open reference mode, OTU quality was similar
between fitting the datasets to reference OTUs with OptiFit and clustering the
datasets _de novo_ with OptiClust. This held true for all datasets and reference
databases. However, in closed reference mode, OTU quality was slightly worse
when fitting to Greengenes and Silva, and much worse when fitting to RDP as
compared to OptiClust. No more than half of query sequences were fit to
reference OTUs in closed reference mode across any dataset/database combination.
**TODO: specify exact numbers for fraction mapped.**
OptiFit was able to fit more query sequences to reference OTUs created with the
Greengenes and Silva databases than with RDP. In terms of run time,
closed reference OptiFit outperformed OptiClust, while OptiClust outperformed
open reference OptiFit.

@@

- **TODO:** for these data, we don't see a compelling reason to use reference-based clustering over _de novo_. it was supposed to speed things up. the reason to do reference-based is if you like the ref OTUs -- e.g. for ML or downstream tools e.g. PiCrust?
- **TODO:** highlight difference between what we're doing and what previously was done.
  others use single ref seq to define an OTU, while we use all the ref & query seqs to define an otu. highlight why ours is better than previous methods.
